% In this file you should put the actual content of the blueprint.
% It will be used both by the web and the print version.
% It should *not* include the \begin{document}
%
% If you want to split the blueprint content into several files then
% the current file can be a simple sequence of \input. Otherwise It
% can start with a \chapter or \chapter for instance.


\newcommand{\docref}[3]{}

\chapter{Introduction}

The aim of this document is to give a detailed proof of Stone duality for Boolean
algebras \cite{Sto1937BA} to facilitate its formalization in Mathlib. Note that
Stone actually proved something more general for bounded distributive lattices
\cite{Sto1937}. (Below, I will cut everything into very small numbered lemmas
and definitions which is not how I usually like to write math, but it seemed
like it might be useful for organizing the formalization. TODO: put this in the
lean blueprint format.)

\chapter{Definitions and statement}
I first recall some definitions from Mathlib. A subset $S$ of a topological
space is \emph{preconnected} if, whenever $U$ and $V$ are open subsets such
that $S \subseteq U \cup V$, $S \cap U \neq \emptyset$ and $S \cap V \neq
\emptyset$, we have $S \cap U \cap V \neq \emptyset$. A topological space is
\emph{totally disconnected} if any preconnected subset is a subsingleton. 
A \emph{Boolean space} is a compact Hausdorff totally disconnected space.%
\footnote{What I call Boolean space is called \texttt{Profinite} in Mathlib and
is sometimes called `Stone space' by other mathematicians.} The category
$\BoolSp$ of Boolean spaces is the full subcategory of topological spaces based
on Boolean spaces.

A \emph{Boolean algebra} is by definition a bounded distributive lattice with
complements. This means that it is a structure of the form $(A, \vee, \wedge,
\zer, \one, \neg)$ satisfying a number of axioms. We do not really care what
the axioms are precisely. What is important to know is: 
\begin{lemma}\lean{Set.instBooleanAlgebraSet}\leanok
    For any set $X$, the structure 
    $(\mathcal{P}(X), \cup, \cap, \emptyset, X, (-)^{\c})$ 
    is a Boolean algebra.
\end{lemma}
In particular, I write $\mbf{2} = \{ \zer, \one \}$ for the unique two-element
Boolean algebra; it can be seen that it is a Boolean algebra because it is the
power set algebra of a singleton set. A Boolean algebra \emph{homomorphism} is
a structure-preserving function between Boolean algebras. In fact, to be a
homomorphism, it suffices to preserve $\vee$, $\zer$, and $\neg$, because the
operations $\one$ and $\wedge$ are term-definable from those. In particular,
for a subset $S$ of a Boolean algebra $A$ to be a subalgebra, it suffices for
$S$ to contain $\zer$, and be closed under the operations $\vee$ and $\neg$. An
\emph{ideal} of a Boolean algebra is a subset $I$ which contains $\zer$ and is
such that $a \vee b \in I$ if, and only if, both $a \in I$ and $b \in I$ (this
is equivalent to the usual definition for rings but formulated in a more
convenient way for Boolean algebras). An ideal is \emph{proper} if it does not
contain $\one$. The \emph{partial order} on a Boolean algebra $A$ can be
defined by $a \leq b$ iff $a \vee b = b$, or equivalently  iff $a \wedge b =
a$. Our aim is to prove:
\begin{theorem}\lean{StoneDuality.Equiv}
    \label{BABoolSpdualeq} 
    The categories $\BoolAlg$ and $\BoolSp$ are dually equivalent.
\end{theorem}
It will also be useful and interesting to actually have concrete definitions of
the dual equivalence functors both ways, as we will do below. 
\begin{remark}
An alternative proof, which is shorter on paper but does not give explicit
access to the definition of the dual equivalence functors, would be to use the
fact that any fully faithful essentially surjective functor is part of an
equivalence (this is \docref{CategoryTheory/Equivalence}
{CategoryTheory.Equivalence.ofFullyFaithfullyEssSurj}{in Mathlib}). However,
the inverse part of the equivalence that is produced by that fact only exists
thanks to an application of axiom of choice, and it seems like it would be hard
to work with. I could of course be wrong about this, and implementing this road
towards the proof might be an interesting alternative experiment. It may have
the advantage of avoiding getting into the weeds of natural transformation
arguments as I need to do in \cref{eta is natural transformation}, \cref{epsilon is natural transformation}, and \cref{triangle} below.
\end{remark}

\chapter{Two useful facts}
We will need the following fundamental fact about Boolean algebras which is
sometimes called the ultrafilter principle. It could also be deduced from the
ring-theoretic fact (probably in Mathlib) that any non-unit element of a ring
is in some maximal ideal, but a direct proof is not hard.
\begin{lemma}\label{ultrafilterprinciple}
    Let $A$ be a Boolean algebra. For any $a \in A \setminus \{\one\}$, 
    there exists a homomorphism $x \colon A \to \mbf{2}$ such that $x(a) =
    \zer$.
\end{lemma}
\begin{proof}
    By Zorn's lemma, let $I$ be a maximal element of the set of proper ideals
    of $A$ which contain $a$. Define $x(b) = \zer$ iff $b \in I$. Clearly,
    $x(a) = \zer$; we need to check that $x$ is a homomorphism. The equalities
    $x(\zer) = \zer$ and $x(b \vee b') = x(b) \vee x(b')$ are easy to check
    from the defining properties of an ideal. To see that $x(\neg b) = \neg
    x(b)$ for any $b \in A$, the crucial observation is that if $b \not\in I$
    and $\neg b \not\in I$, then it is possible to enlarge $I$ by adding $b$ to
    it and generating an ideal $I'$, and $I'$ will still be proper because
    $\neg b \not\in I$. By maximality of $I$ this is impossible.  We thus get
    that, for any $b \in A$, one of $b$ and $\neg b$ must be in $I$, and they
    can never be both in $I$, since that would give $\one = b \vee \neg b$ in
    $I$, again contradicting that $I$ is proper. It now follows from the
    definitions that $x(\neg b) = \neg x(b)$.
\end{proof}

Recall that a space $X$ is called \emph{totally separated} if, for any distinct $x,y \in X$, there exist disjoint open sets $U, V \subseteq X$ such that $x \in U$ and $y \in V$.
\begin{lemma}\lean{instTotallySeparatedSpace} \leanok \label{KTDimpliesTotSep}
    Any compact Hausdorff totally disconnected space is totally separated.
\end{lemma}


\begin{lemma}\lean{isTotallyDisconnected_of_isClopen_set}\leanok
    Suppose that, for any distinct $x,y \in X$, there exists a clopen set $K \subseteq X$ such that $x \in K$ and $y \not\in K$. Then $X$ is totally disconnected.
\end{lemma}

The following equivalent definition of Boolean spaces is convenient. 

\begin{proposition}\label{BoolSpchar}
    A compact Hausdorff space $X$ is totally disconnected if, and only if, for
    any distinct $x, y \in X$, there exists a clopen set $K \subseteq X$ such
    that $x \in K$ and $y \not\in K$.
\end{proposition}

\begin{proof}
    Finally, in any totally separated space, one may actually
   separate distinct points by clopens, and
   \docref{Topology/Connected/TotallyDisconnected}
   {exists_isClopen_of_totally_separated}{this is also in Mathlib}.
    % The necessity is for example proved as the implication (i) $\Rightarrow$
   % (ii) of \cite[Thm~II.4.2]{Joh1986} but I think the following (extracted
   % from there) is a bit simpler, given what is already in Mathlib. Suppose
   % that $X$ is a totally disconnected compact Hausdorff space. It suffices to
   % prove that $X$ is totally separated,
    % because it is in Mathlib that
   % \docref{Topology/Connected/TotallyDisconnected}{exists_isClopen_of_totally_separated}{the
   % separating set in any totally separated space can be chosen to be clopen}.
   % Let $x, y \in X$ be distinct. By total disconnectedness, since $\{x,y\}$
   % is not connected,
    % \docref{Topology/Connected/Basic}{isPreconnected_closed_iff}{we can
   % pick} closed sets $F_1, F_2$ such that $\{x,y\} \subseteq F_1 \cup F_2$
   % and each $F_i$ has non-empty intersection with $\{x,y\}$, but $\{x,y\}
   % \cap F_1 \cap F_2 = \emptyset$. Without loss of generality, we have $x \in
   % F_1$, and then $x \not\in F_2$ (because $x$ is not in $F_1 \cap F_2)$, and
    % therefore (because $F_2$ has non-empty intersection with $\{x,y\}$), we
   % have  $y \in F_2$, so that $y \not\in F_1$ (because $y$ is not in $F_1
   % \cap F_2$). Since $X$ is compact Hausdorff, it is regular, and we can pick
   % disjoint open sets $U$ and $V$ such that $F_1 \subseteq U$, $y \in V$. In
   % particular, $x \in U$, as required.
\end{proof}
\chapter{The functors}
We define the two functors that make up Stone duality for Boolean algebras:
$\Clp \colon \BoolSp \to \BoolAlg$ and $\Spec \colon \BoolAlg \to \BoolSp$.

\paragraph{From spaces to algebras.}
If $X$ is a Boolean space, write $\Clp(X)$ for the set of clopen
subsets of $X$. 

\begin{proposition}\label{clpwelldef}\lean{TopologicalSpace.Clopens.instBooleanAlgebraClopens}\leanok
    $\Clp(X)$ is a Boolean subalgebra of the Boolean algebra $\mathcal{P}(X)$.
\end{proposition}
\begin{proof}
    Finite unions and complements of clopen sets are clopen, and the empty set
    is clopen.
\end{proof} 
\begin{proposition}\label{clpmorphism}\lean{StoneDuality.Clp}\leanok
    If $f \colon X \to Y$ is a continuous function between Boolean spaces, then
    $f^{-1} \colon \Clp(Y) \to \Clp(X)$ is a Boolean algebra homomorphism.
\end{proposition}
\begin{proof}
    $f^{-1}(\emptyset) = \emptyset$, $f^{-1}(K^\c) = f^{-1}(K)^\c$, and
    $f^{-1}(K_1 \cup K_2) = f^{-1}(K_1) \cup f^{-1}(K_2)$ for any clopens $K,
    K_1, K_2$ of $Y$. 
\end{proof}
One sometimes writes $\Clp(f)$ for $f^{-1}$ but the notation is a bit heavy.
\begin{proposition}
    The assignments $X \mapsto \Clp(X)$ and $f \mapsto f^{-1}$ give a
contravariant functor $\BoolSp \to \BoolAlg$.
\end{proposition}
\begin{proof}
    The assignments are well-defined by \cref{clpwelldef} and
    \cref{clpmorphism}. It is a contravariant functor because
    $\ms{id}_X^{-1}(K) = K$ and $(f \circ g)^{-1}(K) = g^{-1}(f^{-1}(K))$ for
    any clopen $K$.
\end{proof}

\paragraph{From algebras to spaces.}
If $A$ is a Boolean algebra, write $\Spec(A)$ for the set of homomorphisms from
$A$ to $\mbf{2}$. We equip $\Spec(A)$ with the subspace topology induced by the
$|A|$-fold power $\mbf{2}^{A}$, where $\mbf{2}$ has the discrete topology.
Define, for any $a \in A$,
\[ U_a \isdef \{ x \in \Spec(A) \ \mid \ x(a) = \one \} \ . \]

\begin{lemma}\label{Uacomp}
    For any $a \in A$, the complement of $U_a$ is equal to $U_{\neg a}$.
\end{lemma}
\begin{proof}
Note that $x \in (U_a)^\c \iff x(a) \neq \one \iff x(a) = \zer \iff x(\neg a) =
\one \iff x \in U_{\neg a} \ .$ 
\end{proof}
\begin{lemma}\label{etahom}
    For any $a, b \in A$ we have $U_a \cup U_b = U_{a \vee b}$, and moreover
    $U_{\zer} = \emptyset$.
\end{lemma}
\begin{proof}
    For any $x \in \Spec(A)$, we have 
    \[ x(a \vee b)= \one \iff x(a) \vee x(b) = \one \iff x(a) = \one \text{ or
    } x(b) = \one \ . \]
    This shows that $U_a \cup U_b = U_{a \vee b}$. To see that $U_{\zer} =
    \emptyset$ just notice that $x(\zer) = \zer \neq \one$.
\end{proof}
\begin{lemma}\label{Uaclp}
    For any $a \in A$, the set $U_a$ is clopen in $\Spec(A)$.
\end{lemma}
\begin{proof}
    For any $a \in A$, the set $U_a$ is open because it is $\pi_a^{-1}(\one)$ where $\pi_a$
    denotes the restriction of the projection function from $\Spec(A)$ to
    $\mbf{2}$ on the coordinate $a$, and this projection function is continuous
    by the definitions of product and subspace topology. The set $U_a$ is
    closed because its complement is equal to $U_{\neg a}$ by \cref{Uacomp},
    and we just showed that any $U_b$ is open.
\end{proof}
\begin{proposition}\label{Specbasis}
The topology on $\Spec(A)$ is generated by the basis
$(U_a)_{a \in A}$ of clopen sets. 
\end{proposition}
\begin{proof}
    In general, the topology on a product $\prod_{i\in I} X_i$ is generated by
    the sets $\pi_i^{-1}(U)$ where $i \in I$ and $U \subseteq X_i$ ranges over
    a basis for the open sets of $X_i$. Now in the case of $\mbf{2}^A$, this
    implies that the topology has as a basis the sets $\pi_a^{-1}(\zer)$ and
    $\pi_a^{-1}(\one)$, as $a$ ranges over the elements of $A$. But
    \cref{Uacomp} gives that $\pi_a^{-1}(\zer) = \pi_{\neg a}^{-1}(\one)$, so
    it actually suffices to take the sets $U_a = \pi_a^{-1}(\one)$ as $a$
    ranges over the elements of $A$.
\end{proof}
\begin{theorem}
    The topological space $\Spec(A)$ is a Boolean space.
\end{theorem}
\begin{proof}
    If $x, y \in \Spec(A)$ are distinct, then there is an element $a \in A$
    such that $x(a) \neq y(a)$. Without loss, say that $x(a) = \one$ and $y(a)
    = \zer$. Then $x \in U_a$ and $y \not\in U_a$. Thus, any two distinct
    elements are separated by a clopen set. This suffices to conclude that
    $\Spec(A)$ is totally disconnected by
    \docref{Topology/Connected/TotallyDisconnected}
    {isTotallyDisconnected_of_isClopen_set}{a result that is in Mathlib
    already}, and it also clearly implies that the space is Hausdorff. It
    remains to prove that $\Spec(A)$ is compact. This is where a weak form of
    axiom of choice must be used. We show that the set of homomorphisms is
    closed as a subspace of $\mbf{2}^{|A|}$, and it is therefore compact, since
    $\mbf{2}^{|A|}$ is compact by Tychonoff's Theorem. In order to see that
    $\Spec(A)$ is closed in $\mbf{2}^{|A|}$, notice that \[ \Spec(A) =
    \bigcap_{a,b\in A} J_{a,b} \cap \bigcap_{a \in A} N_a \cap Z \ , \]
    where $J_{a,b} \isdef \{ x \in \mbf{2}^A \ : \ x(a \vee b) = x(a) \vee x(b)
    \},$ $N_a \isdef  \{x \in \mbf{2}^A \ : \ x(\neg a) = \neg x(a) \},$ $Z
    \isdef \{ x \in \mbf{2}^A \ : \ x(\zer) = \zer\}.$ Each of these sets is
    clopen because its definition only depends on a finite number of
    coordinates of $x$. For example, to spell this out a bit more, for any $a,
    b \in A$,
    \[ J_{a,b} = 
        (\pi_{a \vee b}^{-1}(\zer) \cap \pi_{a}^{-1}(\zer) \cap
        \pi_{b}^{-1}(\zer)) \cup (\pi_{a \vee b}^{-1}(\one) \cap
    (\pi_a^{-1}(\one) \cup \pi_b^{-1}(\one))) \ , \]
    where $\pi_c \colon \mbf{2}^A \to \mbf{2}$ is the projection onto the $c$
    coordinate.\qedhere

    \old{Here is a more concrete proof. By a well-known fact of topology that is
    hopefully in Mathlib, it suffices to prove the finite subcovering property
    with respect to the basis.  Let $T \subseteq A$ be a collection of elements
    with the property that $\bigcup_{a \in T} U_a = \Spec(A)$. Let $I$ be the
    ideal of $A$ generated by $T$. Here ideal means the same as in ring theory,
    but a more convenient definition is the following. This ideal can be
    described concretely as $$I = \{ b \in A \ \mid \ \text{ there is a finite
    } F \subseteq T \text{ such that } b \leq \bigvee F\}\ . $$ If we can prove
    that $\one \in I$, then we are done, because we have a finite $F \subseteq
    T$ such that $\bigvee F = \one$, and then, for any $x \in \Spec(A)$, we get
    \[ \one = x(\one) = x(\bigvee F) = \bigvee_{a \in F} x(a) \, \] so that
    $x(a) = \one$ for some $a \in F$.

    Towards a contradiction, suppose that $\one\not\in I$. By an application of
    Zorn's Lemma, there is a maximal ideal $J$ containing $I$ such that still
    $\one\not\in J$. Let $x \isdef \chi_J \colon A \to \mbf{2}$ be the
    characteristic function of $J$, that is, $x(a) = \zer$ if $a \in J$ and
    $x(a) = \one$ otherwise. One can now prove that $x$ is a homomorphism, as
    follows. By definition $\zer \in I$ so $\zer \in J$, so $x(\zer) = \zer$.
    Also, in any ideal, we have that $a \vee b \in I$ if, and only if, $a \in
    I$ and $b \in I$. This shows that $x(a \vee b) = x(a) \vee x(b)$. Finally,
    we need to show that $x$ preserves negation. Suppose that $x(a) = \one$, so
    that $a \not\in J$. We need to show that $\neg a \in J$. If $\neg a \not\in
    J$, then $\one$ is not in the ideal generated by $J \cup \{a\}$. But this
    ideal strictly contains $J$, contradicting the maximality of $J$. Thus, we
    must have $x(\neg a) = \zer$. Finally, if $x(a) = \zer$, then $a \in J$,
    and we cannot also have $\neg a \in J$, because that would imply $\one = a \vee
    \neg a$ is in $J$ as well, contradiction. Thus, $x(\neg a) = \one$.}
\end{proof}
\begin{definition}
    Given a homomorphism $f \colon A \to B$ we define $f_* \colon \Spec(B) \to
    \Spec(A)$ as the map that sends $y \colon B \to \mbf{2}$ to $y \circ f
    \colon A \to \mbf{2}$. 
\end{definition}
This is well-defined: Since $y \circ f$ is the composition of two
homomorphisms, it is again a homomorphism, showing that $f_*$ is a well-defined
function. 

\begin{proposition}
    For any Boolean algebra homomorphism $f \colon A \to B$, $f_* \colon
    \Spec(B) \to \Spec(A)$ is a continuous function.
\end{proposition}
\begin{proof}
    For any $a \in A$ and $y \in \Spec(B)$, we have
\[ y \in f_*^{-1}(U_a) \iff f_*(y) \in U_a \iff y(f(a)) = \one \iff y \in
U_{f(a)}\ , \]
    so that the inverse image $f_*^{-1}(U_a)$ of any basic open set is again a
(basic) open.
\end{proof}

\begin{proposition}
    The assignment $f \mapsto f_*$ is contravariant functorial from $\BoolAlg$
    to $\BoolSp$. 
\end{proposition}
\begin{proof}
    If $f \colon A \to B$ and $g \colon B
\to C$ then for any $y \in \Spec(C)$,
\[ (\id_C)_*(y) = y \circ \id_C = y, \quad (g \circ f)_*(y) = y \circ g \circ f =
f_*(g_*(y)) \ . \qedhere \]
\end{proof}

One sometimes writes $\Spec f$ for $f_*$ but
the notation is a bit heavy.

\chapter{The equivalence}
\begin{definition}
Let $A$ be a Boolean algebra. We write $\eta_A$ for the Boolean homomorphism
that sends $a \in A$ to $U_a \in \Clp(\Spec(A))$.
\end{definition}
It is indeed a homomorphism by \cref{Uacomp} and \cref{etahom}.
\begin{proposition}
The function $\eta_A$ is bijective, and thus an isomorphism in $\BoolAlg$.
\end{proposition}
\begin{proof}
    The function $\eta_A$ is injective: 
    \cref{ultrafilterprinciple} gives that if $U_a = \Spec(A)$ then 
    $a = 1$. Now if $a \neq b$ then $(\neg a \vee b) \wedge (\neg b \vee a)
    \neq 1$, from which one gets that $\eta(a) \neq \eta(b)$ (this is the
    Boolean algebra version of the usual argument that a ring homomorphism is
    injective if it has zero kernel). 

    The function $\eta_A$ is surjective: Let $K \subseteq \Spec(A)$ be clopen.
    By \cref{Specbasis}, since $K$ is open, there exists $I \subseteq A$ such
    that $K = \bigcup_{a \in I} U_a$. Since $K$ is closed and $\Spec(A)$ is
    compact, $K$ is compact. Pick a finite subset $F \subseteq I$ such that $K
    = \bigcup_{a \in F} U_a$. By \cref{etahom}, we get $K = U_{a_0} =
    \eta_A(a_0)$, where $a_0 \isdef \bigvee F$.
\end{proof}

\begin{definition}
    Let $X$ be a Boolean space. We write $\epsilon_X$ for the function $X \to
    \Spec (\Clp(X))$ that sends $x \in X$ to the homomorphism $\epsilon_X(x)
    \colon \Clp(X) \to \mbf{2}$ defined by sending any $K \in \Clp(X)$ to
    $\one$ if $x \in K$, and to $\zer$ otherwise.
\end{definition}
\begin{lemma}\label{epseq}
    Let $X$ be a Boolean space. For any $K \in \Clp(X)$, we have
    $\epsilon_X^{-1}(U_K) = K$.
\end{lemma}
\begin{proof}
    Note that, for any $x \in X$,
    \[ x \in \epsilon_X^{-1}(U_K) \iff \epsilon_X(x) \in U_K \iff
    \epsilon_X(x)(K) = \one \iff x \in K \ . \qedhere \] 
\end{proof}
\begin{proposition}
    The function $\epsilon_X$ is a homeomorphism, for any Boolean space $X$.
\end{proposition}
\begin{proof}
    $\epsilon_X$ is continuous: By \cref{Specbasis} and a general fact about
    continuous functions, it suffices to check that $\epsilon_X^{-1}(U_K)$ is
    open for every $K \in \Clp(X)$.  This set is equal to $K$ by \cref{epseq},
    and thus open.

    $\epsilon_X$ is injective: For any $x,y \in X$, if $x \neq y$, then there
    is $K \in \Clp(X)$ such that $x \in K$ and $y \not\in K$, using the
    characterization of Boolean spaces given in \cref{BoolSpchar}. Thus,
    $\epsilon_X(x)(K) = \one$ while $\epsilon_X(x)(K) = \zer$.

    $\epsilon_X$ is surjective: Let $\xi \in \Spec(\Clp(X))$. Consider $F
    \isdef \xi^{-1}(\one) = \{ K \in \Clp(X) \ \mid \ \xi(K) = \one \}$. Note
    that, for any finite collection $\{K_1,\dots,K_n\} \subseteq F$, we also
    have $\bigcap_{i=1}^n K_i \in F$, and in particular the intersection is
    non-empty, since $\emptyset \not\in F$. Thus, since $X$ is compact and $F$
    is a filter of closed(-and-open) sets which does not contain $\emptyset$,
    the intersection $\bigcap F$ is non-empty. Pick $x \in \bigcap F$. We claim
    that $\epsilon_X(x) = \xi$. Let $K \in \Clp(X)$ be arbitrary, we need to
    show that $\epsilon_X(x)(K) = \xi(K)$. There are two cases. If $K \in F$,
    then $x \in K$, so $\epsilon_X(x)(K) = \one = \xi(K)$. Otherwise, $K
    \not\in F$. Then $\xi(K) = \zer$, and, using that $\xi$ is a homomorphism,
    we have $\xi(K^\c) = \neg \xi(K) = \one$. Thus, $K^\c \in F$, from which we
    get that $x \in K^\c$, so $\epsilon_X(x)(K) = \zer$, as required.

    Now $\epsilon_X$ is a homeomorphism because it is a continuous bijection
    from a compact to a Hausdorff space.
\end{proof}

\begin{proposition}\label{eta is natural transformation}\lean{StoneDuality.eta}
    The assignment $A \mapsto \eta_A$ is a natural transformation from the
    identity functor on $\BoolAlg$ to the functor $\Clp \circ \Spec$.
\end{proposition}
\begin{proof}
    Let $h \colon A \to B$ be a homomorphism and $a \in A$. For any $x \in
    \Spec(B)$, we have
    \[ x \in \eta_B(h(a)) \iff x(h(a)) = \one \iff x \circ h \in \eta_A(a) \iff
    x \in (\Clp \circ \Spec)(h)(\eta_A(a)) \ , \]
    where the last equivalence follows from the definitions of $\Clp$ and
    $\Spec$ on morphisms. We conclude that $\eta_B(h(a)) = (\Clp \circ
    \Spec)(h)(\eta_A(a))$, as required.
\end{proof}

\begin{proposition}\label{epsilon is natural transformation}\lean{StoneDuality.epsilon}
    The assignment $X \mapsto \epsilon_X$ is a natural transformation from the
    identity functor on $\BoolSp$ to the functor $\Spec \circ \Clp$.
\end{proposition}
\begin{proof}
    Let $f \colon X \to Y$ be a continuous function and $x \in X$. Write $\xi_1
    \isdef \epsilon_Y(f(x))$ and $\xi_2 \isdef ((\Spec \circ
    \Clp)(f))(\epsilon_X(x))$. For any $K \in \Clp(Y)$, we have
    \[ \xi_1(K) = \one \iff f(x) \in K \iff x \in f^{-1}(K) \iff
    \xi_2(K) = \one \ , \]
    where the last equivalence again follows from the definitions of $\Spec$
    and $\Clp$ on morphisms. Thus, $\xi_1 = \xi_2$.
\end{proof}

\begin{remark}\lean{CategoryTheory.Equivalence}\leanok
According to the definition of categorical equivalence in Mathlib, we finally need to prove one
of the triangle laws for the natural isomorphisms $\eta$ and $\epsilon$, namely
the one that says that the composite natural transformation $\Spec \Rightarrow
\Spec \Clp \Spec \Rightarrow \Spec$ is the identity.
\end{remark}
\begin{proposition}\label{triangle}
    The triangle law for $\eta$ and $\epsilon$ holds. 
\end{proposition}
\begin{proof}
    Let $A$ be a Boolean algebra. We write $f \isdef \Spec \eta_A \colon \Spec
    \Clp \Spec A \to \Spec A$ and $g \isdef \epsilon_{\Spec A} \colon \Spec A
    \to \Spec \Clp \Spec A$; we need to prove that $f \circ g$ is the
    identity.\footnote{
        Note that I had to reverse direction of morphisms with respect to what
        is written in the mathlib documentation, because $\Spec$ and $\Clp$ are
        contravariant functors.}
    Let $x \in \Spec A$. We want to show that $x = (f \circ g)(x)$.  Write $y
    \isdef g(x)$. By definition of $g$, we have, for any $K \in \Clp \Spec A$,
    that $y(K) = \one \iff x \in K$. By definition of $f$, we have, for any $a
    \in A$, that $f(y)(a) = \one \iff y(\eta_A(a)) = \one$. Combining these two
    facts, we see that, for any $a \in A$, $f(y)(a) = 1 \iff x \in \eta_A(a)
    \iff x(a) = \one$, by definition of $\eta_A$. Thus, $f(g(x)) = f(y) = x$.
\end{proof}


